Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-08-05 Final_Manuscript.Rmd
more citations

@robertsfredrickson Objectification Theory, suggests that women are objectified in interpersonal encounters, and that the effect might be strongest from men, but possibly also from women. Psychological researchers studying the sexual objectification of women, and the resulting self-objectification that follows, have in recent years explored the interpersonal process of the objectification - self-objectification link (CITE INTERPERSONAL PROCESS LINKS LITERATURE). For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men's reported objectification of their female interaction partner during an interaction was associated with increased self-objectification reported by their partner. @gervais2020social has reviewed the literature on interpersonal self-objectification and has organized the the process in a theoretical model called the SIMO, but this model is only relevant for mixed-gender encounters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there is currently no studies investigating the process of interpersonal objectification same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? For example, perhaps women could be amplifying each others' self-objectification as they get ready to go out for the night, *before* the theoretical encounters with men outlined in @gervais2020social occur.

@@

##Women Objectifying Women
Past research has focused on men’s objectifying behaviors, with little attention to the role that women may also play in enacting objectification of women. @hill2008examining assessed women’s experiences of objectification from men separately from women’s experiences of objectification from women and found that women may be socialized not only to see themselves as objects but perhaps to see other women as objects as well (THEY DIDN'T DETERMINE THIS IN PERSON).
Some studies have found that women objectify other women to a greater extent than they objectify men [@loughnan2015exploring]. Women objectify (dehumanized) sexualized targets presented as images [@puvia2013being], but we know next to nothing about interpersonal sexual objectification among women and what the immediate effects might be. Further, do women objectify non-sexualized women they are interacting with? What might be different for same-gender interactions? @harsey2020men found that self-objectification was related to the objectification of women to a similar degree for men and women participants, although men scored high than women on a measure of belief that objectification of women is natural for men. ASK RANDI-WE KNOW OBJECTIFICATION HAS MORE ADVERSE CONSEQUENCES FOR WOMEN THAN MEN (Gervais, Vescio, & Allen, 2011; Moradi & Huang, 2008; Saguy, Quinn, Dovidio, & Pratto, 2010)...MENTION THAT...HOWEVER DOES WOMEN OBJECT WOMEN TO A GREATER EXTENT THAN MEN MATTER MORE?...STILL TRYING TO FIGURE THAT OUT? ADDITIONALLY @strelan2005women FOUND THAT THE MORE WOMEN SELF-OBJECTIFY, THE MORE THEY OBJECTIFY OTHER WOMEN.

@@

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the literature on stigmatized-stigmatizer interactions (CITE LITERATURE DIDN'T COME ACROSS THIS), viewing the experience of being objectified in an interaction as an identity threat situation (MAYBE CITE MILES-MCLEAN PAPER "STOP LOOKING AT ME!"). When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor]. I DONT KNOW WHERE THIS PAPER IS? ASK RANDI ABOUT IDENTITY_THREAT PAPERS

@@

#Discussion
The current study tested whether the model of interpersonal objectification and state self-objectification (SSO) used in @garcia2016objectification replicates in a same-gender dyadic sample of women only. Although past reserach has found that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], this is the first study to test if *interpersonal* other-objectification by women is related to state self-objectification in their woman-identified interaction partners. As hypothesized, the current study did find a significant relationship bewteen a woman's partner's report of having objectified her and her own post-interaction feelings of self-objectification. That is, there was a significant partner effect of other-objectification on SSO. This effect extends the equivalent relationship found in mixed-gender interactions to the context of same-gender interactions between women. Thus, it could be that it is not only the male gaze, and male other-objectification that is related to women's state self-objectification, but being objectified by another woman can also results in SSO, at least in the context of a scenario where they know they are being evaluated as a potential dating partner.

As is the case in all correlational studies, we cannot be sure in the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation in theoreticall interesting given that, in both the @garcia2016objectification study and in the current study, it is *women's* SSO this relates to being objectified by one's partner. However, Objectification Theory [@robertsfredrickson], as well as past experimental studies [@saguyetal2010; CITE], suggest that the causal flow is from other-objectification to SSO. Although, importantly, the current study found evidence for the partner effect of other-objectification on SSO, there was a lack of evidence for the connections between SSO and downstream negative consequences.

@@

Although we did not find a connection bewteen SSO and authenticity we did find significant positive relationships between authenticity and relationship agency and career aspirations. The relationship between authenticity and cognitive functioning was also estimated as positive, but it was not statistically signfiicant. Again, due to the lack of connection between SSO and authenticity, we found no evidence of *indirect* relationships between SSO and these outcome variables. This evidence of a relationship between authenticity and the relational outcome variables (i.e., relationship agency and career aspirations) provides evidence that corrborates past findings that felt authenticity in interactions is important for healthy relationship functioning [@garcia2016objectification; PWQ paper I reviewed]. Just as authenticity has been found to be important in intergroup interactions [CITE; CITE; CITE], we again find more evidence here that distuptions in feelings of authenticity can negatively impact relationships beyond the current partner. Although the current study did not find a connection between authenticity and SSO, this seems theoretically to be a natural connection, and more work needs to be done to discover when and how SSO leads to inauthenticity in interpersonal objectification situations.

>>>>

@robertsfredrickson Objectification Theory, suggests that women are objectified in interpersonal encounters, and that the effect might be strongest from men, but possibly also from women. Psychological researchers studying the sexual objectification of women, and the resulting self-objectification that follows, have in recent years explored the interpersonal process of the objectification - self-objectification link [@garcia2016objectification, @strelan2018birds, @riemer2020self]. For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men's reported objectification of their female interaction partner during an interaction was associated with increased self-objectification reported by their partner. @gervais2020social has reviewed the literature on interpersonal self-objectification and has organized the the process in a theoretical model called the SIMO, but this model is only relevant for mixed-gender encounters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there is currently no studies investigating the process of interpersonal objectification same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? For example, perhaps women could be amplifying each others' self-objectification as they get ready to go out for the night, *before* the theoretical encounters with men outlined in @gervais2020social occur.

@@

##Women Objectifying Women
Past research has focused on men’s objectifying behaviors, with little attention to the role that women may also play in enacting objectification of women. @hill2008examining assessed women’s experiences of objectification from men separately from women’s experiences of objectification from women and found that women may be socialized not only to see themselves as objects but perhaps to see other women as objects as well (THEY DIDN'T DETERMINE THIS IN PERSON).
Some studies have found that women objectify other women to a greater extent than they objectify men [@loughnan2015exploring]. Women objectify (dehumanized) sexualized targets presented as images [@puvia2013being], but we know next to nothing about interpersonal sexual objectification among women and what the immediate effects might be. Further, do women objectify non-sexualized women they are interacting with? What might be different for same-gender interactions? @harsey2020men found that self-objectification was related to the objectification of women to a similar degree for men and women participants, although men scored high than women on a measure of belief that objectification of women is natural for men. Additionally, @strelan2005women found that the more women self-objectify, the more they objectify other women.  ASK RANDI-WE KNOW OBJECTIFICATION HAS MORE ADVERSE CONSEQUENCES FOR WOMEN THAN MEN (Gervais, Vescio, & Allen, 2011; Moradi & Huang, 2008; Saguy, Quinn, Dovidio, & Pratto, 2010)...MENTION THAT...HOWEVER DOES WOMEN OBJECT WOMEN TO A GREATER EXTENT THAN MEN MATTER MORE?...STILL TRYING TO FIGURE THAT OUT? ADDITIONALLY

@@

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the literature on stigmatized-stigmatizer interactions [@hebl2005promoting] (CITE LITERATURE DIDN'T COME ACROSS THIS), viewing the experience of being objectified in an interaction as an identity threat situation [@miles2015stop]. When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? Woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor].

@@

#Discussion
The current study tested whether the model of interpersonal objectification and state self-objectification (SSO) used in @garcia2016objectification replicates in a same-gender dyadic sample of women only. Although past research has found that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], this is the first study to test if *interpersonal* other-objectification by women is related to state self-objectification in their woman-identified interaction partners. As hypothesized, the current study did find a significant relationship between a woman's partner's report of having objectified her and her own post-interaction feelings of self-objectification. That is, there was a significant partner effect of other-objectification on SSO. This effect extends the equivalent relationship found in mixed-gender interactions to the context of same-gender interactions between women. Thus, it could be that it is not only the male gaze, and male other-objectification that is related to women's state self-objectification, but being objectified by another woman can also results in SSO, at least in the context of a scenario where they know they are being evaluated as a potential dating partner.

As is the case in all correlational studies, we cannot be sure in the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation is theoretically interesting given that, in both the @garcia2016objectification study and in the current study, it is *women's* SSO this relates to being objectified by one's partner. However, Objectification Theory [@robertsfredrickson], as well as past experimental studies [@saguyetal2010; CITE], suggest that the causal flow is from other-objectification to SSO. Although, importantly, the current study found evidence for the partner effect of other-objectification on SSO, there was a lack of evidence for the connections between SSO and downstream negative consequences.

@@

Although we did not find a connection between SSO and authenticity we did find significant positive relationships between authenticity and relationship agency and career aspirations. The relationship between authenticity and cognitive functioning was also estimated as positive, but it was not statistically signficant. Again, due to the lack of connection between SSO and authenticity, we found no evidence of *indirect* relationships between SSO and these outcome variables. This evidence of a relationship between authenticity and the relational outcome variables (i.e., relationship agency and career aspirations) provides evidence that corroborates past findings that felt authenticity in interactions is important for healthy relationship functioning [@garcia2016objectification; PWQ paper I reviewed]. Just as authenticity has been found to be important in intergroup interactions [@brunelletal2010; @garcia2016objectification; @tolman2006looking], we again find more evidence here that distuptions in feelings of authenticity can negatively impact relationships beyond the current partner. Although the current study did not find a connection between authenticity and SSO, this seems theoretically to be a natural connection, and more work needs to be done to discover when and how SSO leads to inauthenticity in interpersonal objectification situations.
